M.Sc. Proposal: \\ \bigskip Computational Electrokinetics

### 2013-03-04 notes/thesis-latex-template/hebrewPart.tex
Update thesis.

בעבודה זו אנו נחקור את התנועה האלקטרוקינטית של חלקיקים בתמיסת אלקטרוליט תחת השפעה של שדה חשמלי,
ונפתח תשתית נומרית לפיתרון איטרטיבי של בעיות פיזיקליות לא-לינאריות מסוג זה.
באמצעות תהליכים אלקטרוקינטיים ניתן להניע ולבצע מניפולציות בחלקיקים זעירים בתחומים שונים, כגון:
ננו-טכנולוגיה, זרימה זעירה, הפרדה וניקוי, התפלת מים, אלקטרופורזה ועוד.

@@

הפותר מאפשר קבלת תוצאות נומריות לתנועת חלקיקים מהסוג שלעיל, בהינתן עוצמת השדה החשמלי המופעל.
התוצאות עבור שדות חלשים אומתו על ידי השוואה לפתרונות אסימפטוטיים של הבעיה.

>>>>

בעבודה זו אנו חוקרים את התנועה האלקטרוקינטית של חלקיקים בתמיסת אלקטרוליט תחת השפעה של שדה חשמלי,
ומפתחים תשתית נומרית לפיתרון איטרטיבי של בעיות פיזיקליות לא-לינאריות מסוג זה.
באמצעות תהליכים אלקטרוקינטיים ניתן להניע ולבצע מניפולציות בחלקיקים זעירים בתחומים שונים, כגון:
ננו-טכנולוגיה, זרימה זעירה, הפרדה וניקוי, התפלת מים, אלקטרופורזה ועוד.

@@

הפותר מאפשר קבלת תוצאות נומריות לתנועת חלקיקים מהסוג שלעיל, בהינתן עוצמת השדה החשמלי המופעל.
התוצאות עבור שדות חלשים אומתו על ידי השוואה לפתרונות אסימפטוטיים של הבעיה.
כמו כן, קיבלנו באמצעות הפותר הנומרי לראשונה פתרונות בשדות חזקים, אשר אין להם עדיין קירובים אסימפטוטיים. מפתרונות אלה ניתן ללמוד על אופי המערכת בשדות חזקים, כולל התנהגות אסימפטוטית של שכבת שפה הנוצרת בחזית החלקיק.

### 2013-03-04 notes/thesis-latex-template/thesis.tex
Update thesis.

\usepackage{graphicx}
\usepackage{tikz}
\usepackage{amsmath} 
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{units}   
\usepackage{float}   
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage[colorlinks]{hyperref}
\usepackage{framed}
\usepackage{url}
\usepackage{longtable}
\usepackage[framed,numbered,autolinebreaks,useliterate]{mcode}
\usepackage{minted}

@@

\begin{figure}
    \begin{center}
    \includegraphics[width=1\textwidth]{figs/quadratic_order.eps}
        \caption[Quadratic convergence of steady-state velocity]{
        The discretization error between the numerical and analytical solutions for 
        $\alpha = 0$, $\Du = 0.5$, $\bar\zeta = 6$ and $R_{max} = 100$. The markers represent
        the discretization errors for various grid sizes, and the dotted lines correspond to linear
        functions of $\beta$, scaled by $4^n$. 
        The graph show that, as the number of grid points in each dimension increases by 2, 
        the discretization error decreases by 4.
        }
	    \label{fig:quadratic}
    \end{center}
\end{figure}

@@

\begin{figure}
    \begin{center}
    \includegraphics[width=.45\textwidth]{figs/LargeBetaV_[128_256_512]_Rmax=10.eps}
    \includegraphics[width=.45\textwidth]{figs/LargeBetaV_[512x512]_Rmax=10_30_100.eps}
        \caption[Large $\beta$ results for $\cU$]
{Large $\beta$ results for $\cU(\beta)$ ($\Du = 1$, $\bar\zeta = 10$, $\alpha = 0.5$). 
The left figure shows the numerical results of $\cU(\beta)$ for $R_{max} = 10$ and the following grid sizes: $128 \times 128$, $256 \times 256$, $512 \times 512$. 
The right figure shows the numerical results of $\cU(\beta)$ for
$512 \times 512$ grid for $R_{max} = 100, 30, 10$. For small values of $\beta$, the linear
regime of $\cU(\beta)$ is dominant. 
For moderate $\beta$ values, the steady state velocity is very close to affine function of $\beta$.
However, for larger values of $\beta$, the numerical solution
deviates significantly from the linear regime, due to convergence problems at $\beta > 6$, 
and for $\beta \sim 7$, the solver fails to converge. The numerical convergence 
depends on the grid resolution near the particle surface -- 
as it becomes finer (when $N_r$ increases or $R_{max}$ decreases), the solver convergence improves.}
	    \label{fig:LargeBetaV_grids}	    
    \end{center}
\end{figure}
\begin{figure}
    \begin{center}
    \includegraphics[width=.45\textwidth]{figs/LargeBetaV_DeltaFromLinear[128_256_512]_Rmax=10.eps}
    \includegraphics[width=.45\textwidth]{figs/LargeBetaV_DeltaFromLinear[512x512]_Rmax=10_30_100.eps}
        \caption[Large $\beta$ results for $\cU$ -- difference from linear regime]
{Large $\beta$ results for $\cU(\beta) - \beta\cU_1$ ($\Du = 1$, $\bar\zeta = 10$, $\alpha = 0.5$ and $\cU_1 \approx 4.258$ is linear regime coefficient). 
The left figure shows the numerical results of $\cU(\beta)$ for $R_{max} = 10$ and the following grid sizes: $128 \times 128$, $256 \times 256$, $512 \times 512$. 
The right figure shows the numerical results of $\cU(\beta)$ for
$512 \times 512$ grid for $R_{max} = 100, 30, 10$.}
	    \label{fig:LargeBetaV_DeltaFromLinear_grids}	    
    \end{center}
\end{figure}
\begin{figure}
    \begin{center}
    \includegraphics[width=.5\textwidth]{figs/LargeBetaRatio.eps}
        \caption[Rate of convergence estimation for $\cU$ at large $\beta$]
        {Rate of convergence estimation for $\cU$ at large $\beta$  
        ($\Du = 1$, $\bar\zeta = 10$, $\alpha = 0.5$), 
        based on the numerical results of $\cU(\beta)$, for the grid sizes of 
        $128 \times 128$, $256 \times 256$, $512 \times 512$. 
        The rate of convergence $r(\beta)$ is estimated by 
        $r(\beta) \triangleq \left(\cU^{4h}(\beta) - \cU^{2h}(\beta)\right)/\left(\cU^{2h}(\beta) - \cU^{h}(\beta)\right)$.
        For small $\beta$, $r(\beta) \rightarrow 4$, showing 
        quadratic convergence for the linear regime. 
        For moderate $\beta$, the convergence rate slows down, 
        and it deteriorates for large $\beta$ values.
        }
	    \label{fig:LargeBetaRatio}
    \end{center}
\end{figure}

@@

\begin{figure}
    \begin{center}
    \includegraphics[width=.45\textwidth]{figs/LargeBeta/sol_beta=1.000e-01_[513x513]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_[1]_Psi.eps}
    \includegraphics[width=.45\textwidth]{figs/LargeBeta/sol_beta=8.000e-01_[513x513]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_[1]_Psi.eps}
    \includegraphics[width=.45\textwidth]{figs/LargeBeta/sol_beta=5.000e+00_[513x513]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_[1]_Psi.eps}
    \includegraphics[width=.45\textwidth]{figs/LinearStokes_Psi.eps}
        \caption[Streamfunction results]
        {$\beta = 0.1, 0.8, 5.0$ and Stokes streamlines results: grid size $512 \times 512$, 
        $R_{max} = 30$, $\Du = 1$, $\bar\zeta = 10$, $\alpha = 0.5$.
        The numerical results for $\Psi$ is scaled by $\beta^{-1}$, so the linear regime
        will appear the same for different $\beta$ values.
        Note that the numerical results show 
        flow pattern very similar to that of the analytical Stokes flow.}
	    \label{fig:LargeBeta_Psi}	    
    \end{center}
\end{figure}

\begin{figure}
    \begin{center}
    \includegraphics[width=.45\textwidth]{figs/PsiMinusUniformFlow/sol_beta=1.000e-01_[513x513]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_[1]_Psi.eps}
    \includegraphics[width=.45\textwidth]{figs/PsiMinusUniformFlow/sol_beta=8.000e-01_[513x513]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_[1]_Psi.eps}
    \includegraphics[width=.45\textwidth]{figs/PsiMinusUniformFlow/sol_beta=5.000e+00_[513x513]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_[1]_Psi.eps}
    \includegraphics[width=.45\textwidth]{figs/PsiMinusUniformFlow/LinearStokes_Psi-uniformflow.eps}
        \caption[Streamfunction results]
        {$\beta = 0.1, 0.8, 5.0$ and Stokes streamlines results after subtraction of the uniform
        flow streamfunction ($\frac{\cU}{2}r^2 \sin^2\theta$): grid size $512 \times 512$, 
        $R_{max} = 30$, $\Du = 1$, $\bar\zeta = 10$, $\alpha = 0.5$.
        The numerical results for $\Psi$ is scaled by $\beta^{-1}$, so the linear regime
        will appear the same for different $\beta$ values.
        Even for the largest beta we could compute, the velocity remains fairly symmetric and 		
        ``uneventful'' also in the framework of the ``laboratory'' rather than that of the particle.}
	    \label{fig:LargeBeta_PsiMinusUniformFlow}	    
    \end{center}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
\section{Discussion} \label{sec:discussion}
A numerical framework for a
nonlinear electrokinetic transport system is described.
The framework is used to construct numerical solvers 
for two electrokinetic problems: the migration of ion-exchangers
and surface-charged inert particle electrophoresis.
The numerical results for weak electric field 
exhibit a good correspondence to the asymptotic models as shown above.
Two interesting phenomena (cubic regime and downstream vortex) were predicted for the ion-exchanger
using the numerical results, and were verified using asymptotic expansion of the nonlinear system.
The electrophoresis problem was solved on a high-resolution grid to validate the asymptotic 
cubic correction of the steady-state velocity, as well as providing insights of the solution behavior 
for moderately strong electric fields.
For strong electric field, the nonlinear effects become dominant and
a boundary-layer structure begins to form.
Due to diminishing ion concentration $C$ at the front of the particle,
a boundary layer of thickness approximately $0.2 e^{-0.5\beta}$ 
is formed with respect to the electric field $\bE$ at the front of the particle.
The value of $E_r / \beta$ at ($r = 1$, $\theta = 0$) is given approximately by 
$e^{\beta/2}$, while the value of $C(r = 1, \theta = 0)$ is given approximately 
by $e^{-\beta/2}$, and their product remains close to 1 for all $\beta$ that we could compute
(as shown in Figure \ref{fig:C_Er_product}).
However, the velocity of the particle remains nearly linear with $\beta$
and the velocity field is very close to being symmetric.

>>>>

\usepackage[format=plain]{caption}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{amsmath} 
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{units}   
\usepackage{float}   
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage[colorlinks]{hyperref}
\usepackage{framed}
\usepackage{url}
\usepackage{longtable}
\usepackage[framed,numbered,autolinebreaks,useliterate]{mcode}
\usepackage{minted}

@@

\begin{figure}
    \begin{center}
    \includegraphics[width=1\textwidth]{figs/quadratic_order.eps}
        \caption[Quadratic convergence of steady-state velocity]{
        The discretization error between the numerical and analytical solutions for 
        $\alpha = 0$, $\Du = 0.5$, $\bar\zeta = 6$ and $R_{max} = 100$. The markers represent
        the discretization errors for various grid sizes, and the dotted lines correspond to linear
        functions of $\beta$, scaled by $4^n$. 
        The graph shows that, as the number of grid points in each dimension increases by 2, 
        the discretization error decreases by 4.
        }
	    \label{fig:quadratic}
    \end{center}
\end{figure}

@@

\begin{figure}
    \begin{center}
    \includegraphics[width=.45\textwidth]{figs/LargeBetaV_[128_256_512]_Rmax=10.eps}
    \includegraphics[width=.45\textwidth]{figs/LargeBetaV_[512x512]_Rmax=10_30_100.eps}
        \caption[Large $\beta$ results for $\cU$]
{Large $\beta$ results for $\cU(\beta)$ ($\Du = 1$, $\bar\zeta = 10$, $\alpha = 0.5$). 
The left figure shows the numerical results of $\cU(\beta)$ for $R_{max} = 10$ and the following grid sizes: $128 \times 128$, $256 \times 256$, $512 \times 512$. 
The right figure shows the numerical results of $\cU(\beta)$ for
$512 \times 512$ grid for $R_{max} = 100, 30, 10$. For small values of $\beta$, the linear
regime of $\cU(\beta)$ is dominant. 
For moderate $\beta$ values, the steady state velocity is very close to an affine function of $\beta$.
However, for larger values of $\beta$, the numerical solution
deviates significantly from the linear regime, due to convergence problems at $\beta > 6$, 
and for $\beta \sim 7$, the solver fails to converge. The numerical convergence 
depends on the grid resolution near the particle surface -- 
as it becomes finer (when $N_r$ increases or $R_{max}$ decreases), the solver convergence improves.}
	    \label{fig:LargeBetaV_grids}	    
    \end{center}
\end{figure}
\begin{figure}
    \begin{center}
    \includegraphics[width=.45\textwidth]{figs/LargeBetaV_DeltaFromLinear[128_256_512]_Rmax=10.eps}
    \includegraphics[width=.45\textwidth]{figs/LargeBetaV_DeltaFromLinear[512x512]_Rmax=10_30_100.eps}
        \caption[Large $\beta$ results for $\cU$ -- difference from linear regime]
{Large $\beta$ results for $\cU(\beta) - \beta\cU_1$ ($\Du = 1$, $\bar\zeta = 10$, $\alpha = 0.5$ and $\cU_1 \approx 4.258$ is the linear regime coefficient). 
The left figure shows the numerical results of $\cU(\beta)$ for $R_{max} = 10$ and the following grid sizes: $128 \times 128$, $256 \times 256$, $512 \times 512$. 
The right figure shows the numerical results of $\cU(\beta)$ for
$512 \times 512$ grid for $R_{max} = 100, 30, 10$.
It is seen that the difference is indeed nearly an affine function of $\beta$ in the large-$\beta$ regime, so long as the numerical solution remains sufficiently accurate.}
	    \label{fig:LargeBetaV_DeltaFromLinear_grids}	    
    \end{center}
\end{figure}
\begin{figure}
    \begin{center}
    \includegraphics[width=.5\textwidth]{figs/LargeBetaRatio.eps}
        \caption[Rate of convergence estimation for $\cU$ at large $\beta$]
        {Rate of convergence estimation for $\cU$ at large $\beta$  
        ($\Du = 1$, $\bar\zeta = 10$, $\alpha = 0.5$), 
        based on the numerical results of $\cU(\beta)$, for the grid sizes of 
        $128 \times 128$, $256 \times 256$, $512 \times 512$. 
        The rate of convergence $r(\beta)$ is estimated by 
        $r(\beta) \triangleq \left(\cU^{4h}(\beta) - \cU^{2h}(\beta)\right)/\left(\cU^{2h}(\beta) - \cU^{h}(\beta)\right)$.
        For small $\beta$, $r(\beta) \rightarrow 4$, showing 
        quadratic convergence for the linear regime. 
        For moderate $\beta$, the convergence rate slows down, 
        and it deteriorates for large $\beta$ values.
        }
	    \label{fig:LargeBetaRatio}
    \end{center}
\end{figure}

@@

\begin{figure}
    \begin{center}
    \includegraphics[width=.45\textwidth]{figs/LargeBeta/sol_beta=1.000e-01_[513x513]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_[1]_Psi.eps}
    \includegraphics[width=.45\textwidth]{figs/LargeBeta/sol_beta=8.000e-01_[513x513]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_[1]_Psi.eps}
    \includegraphics[width=.45\textwidth]{figs/LargeBeta/sol_beta=5.000e+00_[513x513]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_[1]_Psi.eps}
    \includegraphics[width=.45\textwidth]{figs/LinearStokes_Psi.eps}
        \caption[Streamfunction results]
        {$\beta = 0.1, 0.8, 5.0$ and Stokes streamlines results: grid size $512 \times 512$, 
        $R_{max} = 30$, $\Du = 1$, $\bar\zeta = 10$, $\alpha = 0.5$.
        The numerical results for $\Psi$ are scaled by $\beta^{-1}$, so the linear regime
        will appear the same for different $\beta$ values.
        Note that the numerical results show a
        flow pattern very similar to that of the analytical Stokes flow.}
	    \label{fig:LargeBeta_Psi}	    
    \end{center}
\end{figure}

\begin{figure}
    \begin{center}
    \includegraphics[width=.45\textwidth]{figs/PsiMinusUniformFlow/sol_beta=1.000e-01_[513x513]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_[1]_Psi.eps}
    \includegraphics[width=.45\textwidth]{figs/PsiMinusUniformFlow/sol_beta=8.000e-01_[513x513]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_[1]_Psi.eps}
    \includegraphics[width=.45\textwidth]{figs/PsiMinusUniformFlow/sol_beta=5.000e+00_[513x513]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_[1]_Psi.eps}
    \includegraphics[width=.45\textwidth]{figs/PsiMinusUniformFlow/LinearStokes_Psi-uniformflow.eps}
        \caption[Streamfunction results]
        {$\beta = 0.1, 0.8, 5.0$ and Stokes streamlines results after subtraction of the uniform
        flow streamfunction ($\frac{\cU}{2}r^2 \sin^2\theta$): grid size $512 \times 512$, 
        $R_{max} = 30$, $\Du = 1$, $\bar\zeta = 10$, $\alpha = 0.5$.
        The numerical results for $\Psi$ are scaled by $\beta^{-1}$, so the linear regime
        will appear the same for different $\beta$ values.
        Even for the largest beta we could compute, the velocity remains fairly symmetric and 		
        ``uneventful'' also in the framework of the ``laboratory'' rather than that of the particle.}
	    \label{fig:LargeBeta_PsiMinusUniformFlow}	    
    \end{center}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
\section{Discussion} \label{sec:discussion}
A numerical framework for a
nonlinear electrokinetic transport system is described.
The framework is used to construct numerical solvers 
for two electrokinetic problems: the migration of ion-exchangers
and surface-charged inert particle electrophoresis.
The numerical results for weak electric field 
exhibit a good correspondence to the asymptotic models as shown above.
Two interesting phenomena (cubic regime and downstream vortex) were predicted for the ion-exchanger
using the numerical results, and were verified using an asymptotic expansion of the nonlinear system.
The electrophoresis problem was solved on a high-resolution grid to validate the asymptotic 
cubic correction of the steady-state velocity, as well as providing insights into
the solution behavior for moderately strong electric fields.
For strong electric field, the nonlinear effects become dominant and
a boundary-layer structure begins to form.
Due to diminishing ion concentration $C$ at the front of the particle,
a boundary layer of thickness approximately $0.2 e^{-0.5\beta}$ 
is formed with respect to the electric field $\bE$ at the front of the particle.
The value of $E_r / \beta$ at ($r = 1$, $\theta = 0$) is given approximately by 
$e^{\beta/2}$, while the value of $C(r = 1, \theta = 0)$ is given approximately 
by $e^{-\beta/2}$, and their product remains close to 1 for all $\beta$ that we could compute
(as shown in Figure \ref{fig:C_Er_product}).
However, the velocity of the particle remains nearly linear with $\beta$
and the velocity field is very close to being symmetric.
